---
title: "penguin flippler lengths anova"
author: "Lily Zhao"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. load packages and read in penguins.csv data
packages: tidyverse, car, onewaytests
```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins<- read_csv("penguins.csv")
```
Is there a significant difference in mean flipper lengths for female chinstrap, gentoo, and adela penguins near Palmer. 

Single factor: species
how many levels exist within that factor: 3
what is my one continous variable: flipper length that we will use to compare means

maybe i can use a one way anova
- assummptions tied to that:
-independent observations
-equal variance which we will use a levenes test for)
- normality assumption (histogram and qq plot)

how many observations exist within each group?
table question will automatically calculates counts based on how many factors you give it
```{r}
counts <- with(penguins, table(species,sex))
View(counts)
#by the clt theorem sample means will be normally distributed but first lets look at the data to see if mean is a good metric of central tendancy to compare. 


penguins2<- penguins %>% 
mutate(
  common_name=
    case_when(
      species=="Chinstrap penguin (Pygoscelis antarctica)"~ "Chinstrap",
      species=="Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
      species=="Adelie Penguin (Pygoscelis adeliae)"~"Adelie"
    )
) %>% 
select(-species) %>% #GETS RID OF THAT OG COLUMN
  filter(sex=="FEMALE")

```

Exploratory graphs and Levene's test for equal variances


```{r}
hists<- ggplot(penguins2, aes(x= flipper_length))+
geom_histogram(aes(fill=common_name))+
  facet_wrap(~common_name, scale="free")
hists

#qqplot to answer questions about normality

#remember to change x to sample=
qq<- ggplot(penguins2, aes(sample= flipper_length))+
geom_qq(aes(fill=common_name))+
  facet_wrap(~common_name, scale="free")
qq

#base on qq and histograms obs for each group appear normally distributed,  also n>30 for all groupps, so menas will be nomrally distributed by CLT. 
```


